PineAPPL: combining EW and QCD corrections\\ for fast evaluation of LHC processes

### 2021-03-26 paper/answer_to_the_editor.txt

Dear Editor,

We would like to thank the referee for carefully reading the manuscript and
making useful comments and suggestions.

All changes are marked in the pre-compiled document (diff.pdf) as
red/strikethrough (removed text) and blue/underlined (added text).

@@

In the following our answers to the issues raised by the referee:

- Issues 1, 3, 4, 5, 6, 8, 9, and 11 have all been addressed as suggested, or
  corrected otherwise.
- Issue 2: The decomposition of the fixed-order expansion written down in eq.
  (2.1) holds to all orders in both the strong and electroweak coupling, and
  therefore the mentioned sentence is correct, as it only concerns PineAPPL.
  The Monte Carlo that we used is limited to NLO, of course.
- Issue 7: For Drell-Yan we checked (compare
  figures/pineappl_ATLASZHIGHMASS49FB.pdf against
  figures/pineappl_ATLASZHIGHMASS49FB_dyn.pdf) that a dynamical scale set to
  Mll does change the NLO QCD in the largest invariant mass bins and therefore
  also the scale uncertainty. However, in comparison to the shown data
  uncertainty the changes are very minor, and therefore we did not update
  plots and text. For the Z transverse momentum process we did update both
  the plot and text, where we indeed see a significant increase of the scale
  variation band in the large transverse momentum bins. The electroweak
  corrections are also smaller, since we normalise to NLO QCD.
- Issue 10: A brief discussion on this matter has been added at the beginning
  of section 3. The reference suggested (Buccioni:2020cfi), is cited in section 
  3.2.1, in the discussion relevant for Drell-Yan.
- Issue 12: Although we agree with the referee that the mentioned tables are
  interesting only to a few readers, the main text refers to it in a few
  places and therefore removing it from the main document would require bigger
  changes. We see no harm in leaving the tables in the document and therefore
  decided to keep them.

### 2021-03-26 paper/sec-doublecounting.tex
Fix errors found by `lacheck`

A first example is the subtraction of (irreducible) background processes which must not be considered as such. A very blatant case
is neutral-current Drell--Yan, where the signal process is the production of an opposite-sign lepton pair, which starts
at $\mathcal O(\alpha^2)$. Because this process is usually thought
as a quark-initiated $s$-channel mechanism ($\mathrm{q} \bar{\mathrm{q}} \to \gamma^*/\mathrm{Z} \to \ell \bar{\ell}$), in many analyses the PI component,
$\gamma \gamma \to \ell \bar{\ell}$ in the $t$ channel, is considered a different process, and therefore as a background and subtracted.
The subtraction from the measured data is done by calculating the theoretical predictions of the double-photon initiated contribution, possibly including (ill-defined) higher-order
corrections. For example, in refs.~\cite{Aaboud:2017ffb,Aad:2016zzw} (a similar statement appears also in an older analysis~\cite{Aad:2013iua}), one reads:
\begin{quote}
The photon-induced process, $\gamma\gamma \to \ell \bar{\ell}$, is simulated at LO using Pythia 8
and the MRST2004qed PDF set~\cite{Martin:2004dh}. The expected yield for this process also accounts for 
NLO QED/EW corrections from references~\cite{Bardin:2012jk,Bondarenko:2013nu}, which decrease the yield by approximately \SI{30}{\percent}.
\end{quote}
Such a distinction, which is unphysical and incorrect in quantum mechanics, may be somehow justified at LO. Beyond this order, it is simply wrong.
Indeed, at $\mathcal O(\alpha^3)$, the reaction $\mathrm{q} \gamma \to \ell \bar{\ell} \mathrm{q}$ becomes possible, which
includes both kind of topologies discussed above, and needs both in order to yield an IR-finite result, see Fig.~\ref{fig:dy-pi} (as a consequence, one cannot speak of EW corrections to $\gamma \gamma \to \ell \bar{\ell}$). It may be useful for the reader to consider a QCD counterpart of this issue, 
e.g. the subtraction of gluon-initiated contribution to top-pair production, in order to get convinced of the incorrectness of this procedure.

>>>>

A first example is the subtraction of (irreducible) background processes which must not be considered as such. A very blatant case
is neutral-current Drell--Yan, where the signal process is the production of an opposite-sign lepton pair, which starts
at $\mathcal O(\alpha^2)$. Because this process is usually thought
as a quark-initiated $s$-channel mechanism ($\mathrm{q} \bar{\mathrm{q}} \to \gamma^*/\mathrm{Z} \to \ell \bar{\ell}$), in many analyses the PI component,
$\gamma \gamma \to \ell \bar{\ell}$ in the $t$ channel, is considered a different process, and therefore as a background and subtracted.
The subtraction from the measured data is done by calculating the theoretical predictions of the double-photon initiated contribution, possibly including (ill-defined) higher-order
corrections. For example, in refs.~\cite{Aaboud:2017ffb,Aad:2016zzw} (a similar statement appears also in an older analysis~\cite{Aad:2013iua}), one reads:
\begin{quote}
The photon-induced process, $\gamma\gamma \to \ell \bar{\ell}$, is simulated at LO using Pythia 8
and the MRST2004qed PDF set~\cite{Martin:2004dh}. The expected yield for this process also accounts for 
NLO QED/EW corrections from references~\cite{Bardin:2012jk,Bondarenko:2013nu}, which decrease the yield by approximately \SI{30}{\percent}.
\end{quote}
Such a distinction, which is unphysical and incorrect in quantum mechanics, may be somehow justified at LO\@. Beyond this order, it is simply wrong.
Indeed, at $\mathcal O(\alpha^3)$, the reaction $\mathrm{q} \gamma \to \ell \bar{\ell} \mathrm{q}$ becomes possible, which
includes both kind of topologies discussed above, and needs both in order to yield an IR-finite result, see Fig.~\ref{fig:dy-pi} (as a consequence, one cannot speak of EW corrections to $\gamma \gamma \to \ell \bar{\ell}$). It may be useful for the reader to consider a QCD counterpart of this issue, 
e.g.\ the subtraction of the gluon-initiated contribution to top-pair production, in order to understand the incorrectness of this procedure.

### 2021-03-26 paper/sec-results.tex

We choose a dynamic value for the renormalisation and factorisation scales $\mu_\mathrm{R}=\mu_\mathrm{F}=\sqrt{M_{\ell \bar{\ell}}^2 + \left( p_\mathrm{T}^{\ell \bar{\ell}} \right)^2}$, used for example in ref.~\cite{Ridder:2016nkl}, where $M_{\ell \bar{\ell}}$ is the invariant mass of the lepton pair and $p_\mathrm{T}^{\ell \bar{\ell}}$ the transverse momentum of it.\footnote{In the last bin the dynamic scale is $\mu_\mathrm{R} = \mu_\mathrm{F} \approx p_\mathrm{T}^{\ell \bar{\ell}}$, and therefore the bin limits drive the scale beyond the limit $Q_\text{max}^2 = (\SI{1}{\tera\electronvolt})^2$ defined in sec.~\ref{sec:grid-representation}, which we increase for this process to $Q_\text{max}^2 = (\SI{10}{\tera\electronvolt})^2$, along with $N_\tau = 40$.}
Consistently with the
experimental analysis, we require $p_\mathrm{T}^\ell>\SI{25}{\giga\electronvolt}$,
$|\eta_\ell|<2.4$, $|M_{\ell\bar\ell} - M_\mathrm{Z}| < \SI{15}{\giga\electronvolt}$,
$|y_{\ell\bar\ell}|<2.4$ and $\SI{20}{\giga\electronvolt}<p_\mathrm{T}^{\ell\bar\ell}<\SI{1500}{\giga\electronvolt}$ for the
transverse momentum and pseudorapidity of each lepton, and for the invariant
mass, pseudorapidity and transverse momentum of the lepton pair.
We finally discard all bins with
$p_\mathrm{T}^{\ell \bar{\ell}}<\SI{20}{\giga\electronvolt}$ to avoid a kinematic
region where resummation effects are sizeable.

>>>>

We choose a dynamic value for the renormalisation and factorisation scales $\mu_\mathrm{R}=\mu_\mathrm{F}=\sqrt{M_{\ell \bar{\ell}}^2 + \left( p_\mathrm{T}^{\ell \bar{\ell}} \right)^2}$, used for example in ref.~\cite{Ridder:2016nkl}, where $M_{\ell \bar{\ell}}$ is the invariant mass of the lepton pair and $p_\mathrm{T}^{\ell \bar{\ell}}$ its transverse momentum.\footnote{In the last bin the dynamic scale is $\mu_\mathrm{R} = \mu_\mathrm{F} \approx p_\mathrm{T}^{\ell \bar{\ell}}$, and therefore the bin limits drive the scale beyond the limit $Q_\text{max}^2 = (\SI{1}{\tera\electronvolt})^2$ defined in sec.~\ref{sec:grid-representation}, which we increase for this process to $Q_\text{max}^2 = (\SI{10}{\tera\electronvolt})^2$, along with $N_\tau = 40$.}
Consistently with the
experimental analysis, we require $p_\mathrm{T}^\ell>\SI{25}{\giga\electronvolt}$,
$|\eta_\ell|<2.4$, $|M_{\ell\bar\ell} - M_\mathrm{Z}| < \SI{15}{\giga\electronvolt}$,
$|y_{\ell\bar\ell}|<2.4$ and $\SI{20}{\giga\electronvolt}<p_\mathrm{T}^{\ell\bar\ell}<\SI{1500}{\giga\electronvolt}$ for the
transverse momentum and pseudorapidity of each lepton, and for the invariant
mass, pseudorapidity and transverse momentum of the lepton pair.
We finally discard all bins with
$p_\mathrm{T}^{\ell \bar{\ell}}<\SI{20}{\giga\electronvolt}$ to avoid a kinematic
region where resummation effects are sizeable.
